---
title: "Data collection"
---

```{r}
library(magrittr)

# Setup
data <- CARS::get_tidy_data_file("2024_data.csv")
config <- yaml::read_yaml("../../config.yml")

data <- CARS::clean_data(data, config)

all_wave_data <- CARS::get_all_waves()

```

:::{.callout-note}
Please note, these are the initial summary statistics for CARS 2024 and further analysis will follow. 
We advise linking directly to this document when distributing to ensure the most up to date information.
:::


# How we collect data

The Coding in Analysis and Research Survey (CARS) data collection takes place for approximately one month, every autumn. The survey is self-selecting and participation is voluntary. Launch dates vary slightly by year to maximise response rate, for example by avoiding clashes with other internal surveys. In 2024, data collection took place from 22 October to 31 December.

We invite analysts to participate in the survey using a variety of online channels, mailing lists, networks and newsletters. For the past four years, the most common source of data has been through departmental Reproducible Analytical Pipeline (RAP) champions, who promote the survey in their organisations. We rely on various champion networks, Heads of Profession (HoPs) for analysis and Departmental Directors of Analysis (DDans) to promote the survey and encourage their analytical communities to participate. This means the response rate and any selection bias will vary across organisations.

Our promotional materials make it clear that we are interested in responses from all analysts, whether or not they use coding in their work. The survey may however attract a disproportionate number of respondents who have an interest in coding and RAP. We advise against making strong inferences about differences between professions and departments or attempting to estimate real frequencies from the data because of these potential limitations.

Lastly, while the survey is open to all public sector analysts, the vast majority of responses come from the UK and devolved Civil Service (`r round(sum(data$workplace == "Civil service, including devolved administrations") / nrow(data) * 100, 1)`% in 2024). As such, follow-up questions on grade and profession applied only to civil servants.

## Where our data comes from

Link tracking allows us to see where responses are coming from. Links promoted by RAP champions were the most commonly used for the past three waves, and accounted for over half of responses in 2024.

```{r}
rename_list <- c(config[["tracking_link"]],
                 list(
                    "HoPs managers support network + GSG Teams Channel" = "Profession newsletters/mailing lists", 
                    "RAP Champions Network" = "RAP Champions",                           
                    "NHS-R community" = "Other",                                 
                    "DATA SCIENCE SLACK" = "Slack",                              
                    "quality teams" = "Other",                                    
                    "HoPs weekly email" = "Profession newsletters/mailing lists",                               
                    "GORS Newsletter" = "Profession newsletters/mailing lists",                                  
                    "DDaT newsletter" = "Profession newsletters/mailing lists",                                 
                    "GSR Friday Bulletin" = "Profession newsletters/mailing lists",                              
                    "AF newsletter" = "Profession newsletters/mailing lists",                                   
                    "ONS SCS"  = "Other",                                         
                    "Reggie" = "Other",                                         
                    "HoP/DDan mailing list" = "Profession newsletters/mailing lists",                            
                    "Economist newsletter" = "Profession newsletters/mailing lists",                            
                    "Geography "  = "Profession newsletters/mailing lists",                                      
                    "Actuaries"   = "Profession newsletters/mailing lists",                                     
                    "RAP champs" = "RAP Champions",                                       
                    "GSS newsletter" = "Profession newsletters/mailing lists",                                  
                    "ONS" = "Other",                                               
                    "data science slack - general" = "Slack",                    
                    "RAS"   = "Other",                                             
                    "Quality champions" = "Other Champions",                               
                    "weekly professions notes" = "Profession newsletters/mailing lists",                        
                    "Analysis Function"  = "Profession newsletters/mailing lists",                              
                    "GORS" = "Profession newsletters/mailing lists",                                             
                    "MQD DASH Group" = "Other",                                   
                    "RAP collaboration slack"  = "Slack",                         
                    "DDAT" = "Profession newsletters/mailing lists",                                            
                    "Final DDANs Extension" = "Profession newsletters/mailing lists",                            
                    "CARS presentation"  = "Other",                               
                    "RAS mailing list" = "Other",                                  
                    "Social research newsletter" = "Profession newsletters/mailing lists",                      
                    "Harmonisation Champions" = "Other Champions",                          
                    "Geography newsletter" = "Profession newsletters/mailing lists",                            
                    "BEIS RAP champs" = "RAP Champions",                                  
                    "HMRC RAP champs" = "RAP Champions",                                 
                    "DSTL RAP champs" = "RAP Champions",                                  
                    "RAP champs round 2" = "RAP Champions",                              
                    "Slack final reminder" = "Slack",                             
                    "RAS round 2"   = "Other",                                    
                    "Evaluation Champions" = "Other Champions",                             
                    "weekly professions note" = "Profession newsletters/mailing lists",                         
                    "HMCLG"= "Other"
                    )
                 ) 
  
all_wave_data$tracking_link <-dplyr::recode(all_wave_data$tracking_link , !!!rename_list)

links <- table(all_wave_data$tracking_link)
most_used <- names(links)[links >= 100]
all_wave_data$tracking_link <- ifelse(!all_wave_data$tracking_link %in% most_used, "Other", all_wave_data$tracking_link)

tracking_link_freqs <- table(all_wave_data$year, all_wave_data$tracking_link) %>% 
  data.frame %>% 
  dplyr::group_by(Var1) %>% 
  dplyr::mutate(Count = sum(Freq), n = Freq / sum(Freq)) %>% 
  dplyr::arrange(Var1) %>%
  data.frame()

order <- rev(order(tracking_link_freqs$n[tracking_link_freqs$Var1=="2024"]))
tracking_link_freqs <- tracking_link_freqs[c(order, order+5, order+10, order+15) ,]
tracking_link_freqs <- dplyr::rename(tracking_link_freqs, "Tracking link" = Var2)

CARS::df_to_table(tracking_link_freqs[c(2,1,5)], 
                  column_headers = c("2021", "2022", "2023", "2024"), 
                  crosstab = T,
                  sample = FALSE)

```

## Sample size by year

```{r}

year_dist <- table(all_wave_data$year) %>% data.frame 
colnames(year_dist) <- c("Year", "Sample")

year_dist %>%
  kableExtra::kable() %>%
  kableExtra::kable_styling()

```

# Respondent characteristics

## Coding frequency

Every year, we ask respondents how often they code to achieve work objectives. While our communication strategy has changed over time, particularly to encourage more non-coders to respond, the findings remain consistent, with a gradual increase in the number of coders over time. Although we seek responses from all analysts the data probably over-represents people with current or prior coding experience.

```{r}
all_wave_data$code_freq <- gsub("All the time", "Always", all_wave_data$code_freq)

all_wave_data$code_freq <- factor(all_wave_data$code_freq, config$code_freq$levels)

code_freqs_by_year <- table(all_wave_data$year, all_wave_data$code_freq) |>
  data.frame() |>
  dplyr::group_by(Var1) |>
  dplyr::mutate(sample = sum(Freq), n = Freq / sum(Freq)) |>
  data.frame() |>
  CARS::get_ci(3, 4)

plot <- CARS::plot_stacked(code_freqs_by_year[c(1,2,5)], 
                           orientation = "v", 
                           type = "bar", 
                           font_size = 14, 
                           xlab = "Year") |>
  plotly::layout(legend = list(traceorder = "reversed"))

table <- CARS::df_to_table(data = code_freqs_by_year[c(2,1, 4, 5)], 
                     crosstab = TRUE, 
                     column_headers = c("In your current role, how often do you write code to complete your work objectives?", 
                                        "2021", "2023", "2023", "2024"))

CARS::wrap_outputs("code-freq", plot, table)

```

## Grade

Across all years, over 80% of Civil Service respondents reported that they are at H, S or Grade 7 grades (or equivalent). While this will be representative of the grade distribution of analysts in some government organisations, it may not be the case for all organisations.

```{r}
all_wave_data$cs_grade <- dplyr::coalesce(all_wave_data$cs_grade, all_wave_data$CS_grade) 

rename_list <- list(
 "Senior Executive Officer (or equivalent)" = "Senior Executive Officer",                                                                    
 "Grade 7 (or equivalent)" = "Grade 7",                                                                                     
 "Grade 6 (or equivalent)" = "Grade 6 or above",                                                                                     
 "Higher Executive Officer (or equivalent)" = "Higher Executive Officer",                                                                    
 "Executive Officer (or equivalent)" = "Administrative Officer or Executive Officer",                                                                           
 "Fast Stream" = "Senior Executive Officer",                                                                                                 
 "SCS Pay Band 1 (or equivalent)" = "Grade 6 or above",                                                                              
 "B1" = "Grade 6 or above",                                                                                                          
 "Administrative Officer (or equivalent)" = "Administrative Officer or Executive Officer",                                                                      
 "~ Skill Zone 4 " = "Administrative Officer or Executive Officer",                                                                                             
 "Deputy Principal Statistician - NI" = "Senior Executive Officer",                                                                          
 "Staff officer equivalent" = "Higher Executive Officer",                                                                                    
 "Assistant Statistician (AS/SO)" = "Higher Executive Officer",                                                                              
 "Deputy Principal Statistician" = "Senior Executive Officer",                                                                               
 "Staff Officer Equivalent (NICS)" = "Higher Executive Officer",                                                                             
 "Apprentice." = "Administrative Officer or Executive Officer",                                                                                                 
 "Deputy Principal" = "Senior Executive Officer",                                                                                            
 "Statistical Officer" = "Higher Executive Officer",                                                                                         
 "Assistant Statistician (Staff Officer Equivalent)" = "Higher Executive Officer",                                                           
 "Deputy Principal Statistician (not sure what equivalent is)" = "Senior Executive Officer",                                                 
 "Staff officer" = "Higher Executive Officer",                                                                                               
 "DP stat" = "Senior Executive Officer",                                                                                                     
 "Staff Officer" = "Higher Executive Officer",                                                                                               
"Cefas Pay Band 5 (not sure of the equivalent)" = "Higher Executive Officer",                                                               
"assistant statistician" = "Higher Executive Officer",                                                                                      
"Assistant Statistician" = "Higher Executive Officer",                                                                                      
"FR PB4" = "Higher Executive Officer",                                                                                                      
"STO" = "Higher Executive Officer",                                                                                                         
"B3" = "Higher Executive Officer",                                                                                                          
"Assistant Statistician, equivalent to Staff Officer (NICS)" = "Higher Executive Officer",                                                  
"Assistant statistician (Staff officer equivalent in Northern Ireland)" = "Higher Executive Officer",                                       
"staff officer in NICS - HEO" = "Higher Executive Officer",                                                                                 
"NHS band 7" = "Senior Executive Officer",                                                                                                  
"Student Placement" = "Administrative Officer or Executive Officer",                                                                                           
"Staff Officer equivalent grade (NICS)  " = "Higher Executive Officer",                                                                     
"DCMS Band B" = "Higher Executive Officer",                                                                                                 
"B2" = "Higher Executive Officer",                                                                                                          
"Forestry Commission PB5" = "Administrative Officer or Executive Officer",                                                                                     
"Unsure - FR Pay Band 5" = "Administrative Officer or Executive Officer",                                                                                      
"EA Grade 5" = "Senior Executive Officer",                                                                                                  
"NHS Band 5 " = "Higher Executive Officer",                                                                                                 
"Grade 5 PHS" = "Higher Executive Officer",                                                                                                 
"Band 5 NHS" = "Higher Executive Officer",                                                                                                  
"Junior Trainee" = "Administrative Officer or Executive Officer",                                                                                              
"Scottish Government: B3" = "Senior Executive Officer",                                                                                     
"Apprentice" = "Administrative Officer or Executive Officer",                                                                                                  
"Statistical Officer (below Assistant Statistician)" = "Higher Executive Officer",                                                          
"Student " = "Administrative Officer or Executive Officer",                                                                                                    
"trainee" = "Administrative Officer or Executive Officer",                                                                                                     
"NHS AfC Band 7" = "Senior Executive Officer",                                                                                              
"Student placement" = "Administrative Officer or Executive Officer",                                                                                           
"B2 Assistant Statistician" = "Higher Executive Officer",                                                                                   
"NHS Band 5" = "Administrative Officer or Executive Officer",                                                                                                  
"SEO+" = "Senior Executive Officer",                                                                                                        
"Student" = "Administrative Officer or Executive Officer",                                                                                                     
"Band B (HEO/SEO equivalent)" = "Higher Executive Officer",                                                                                 
"Grade B3 in SG" = "Senior Executive Officer",                                                                                              
"B3 - TRS to C1" = "Senior Executive Officer",                                                                                              
"Scot Gov B3" = "Senior Executive Officer",                                                                                                 
"Band B (HEO + SEO)" = "Higher Executive Officer",                                                                                          
"Research Officer" = "Higher Executive Officer",                                                                                            
"Trainee " = "Administrative Officer or Executive Officer",                                                                                                    
"Scottish Government B3 TRS" = "Senior Executive Officer",                                                                                  
"seconded Chief Scientific Adviser" = "Grade 6 or above",                                                                           
"Junior Actuary Trainee " = "Administrative Officer or Executive Officer",                                                                                     
"Scottish Government Band B2" = "Higher Executive Officer")  

all_wave_data$cs_grade <- dplyr::recode(all_wave_data$cs_grade , !!!rename_list)


all_wave_data$cs_grade <- factor(all_wave_data$cs_grade, levels = c(
  "Administrative Officer or Executive Officer",
  "Higher Executive Officer",
  "Senior Executive Officer",
  "Grade 7",
  "Grade 6 or above"
))

grade_freqs <- table(all_wave_data$year, all_wave_data$cs_grade) %>% data.frame() %>%
  dplyr::group_by(Var1) %>%
  dplyr::reframe(Var2, n= Freq / sum(Freq)) %>%
  data.frame

plot <- CARS::plot_stacked(grade_freqs, orientation = "v", font_size = 14, xlab = "Year", colour_scale = "gradient") %>%
  plotly::layout(legend = list(traceorder = "reversed"))

table <-  CARS::df_to_table(grade_freqs[c(2,1,3)], crosstab = T, column_headers = c("Grade", "2020", "2021", "2022", "2023"))

CARS::wrap_outputs("grades-by-year", plot, table)
```

## Profession

Below is a breakdown of the proportion of respondents in different Civil Service professions. These cover the [Analysis Function professions](https://analysisfunction.civilservice.gov.uk/about-us/frequently-asked-questions/) and do not apply outside of the civil service. The exception to these are data scientists and data engineers who do not have an official government profession. They are included separately here to avoid skewing the data for other professions. Note that respondents can be members of more than one analytical profession. Profession data is difficult to compare across years as these questions have changed in line with changes to the Analysis Function.

The CARS sample has high representation from statisticians compared with other professions. This again may be representative of some organisations but not all.

```{r}
prof_cols <- colnames(data)[grepl("prof_", colnames(data))]
prof_data <- data.frame(data[prof_cols] != "No")

frequencies <- colSums(prof_data, na.rm = T) |> data.frame() %>% tibble::rownames_to_column()
colnames(frequencies) <- c("Profession", "Count")
frequencies$n <- frequencies$Count / nrow(data)
frequencies <- dplyr::arrange(frequencies, dplyr::desc(n))

frequencies$Profession <- dplyr::recode(frequencies$Profession, !!!config[["professions"]][["cols"]])

CARS::df_to_table(frequencies[c(1,3)],
                  column_headers = c("Civil Service Profession", 
                                              "Percentage"),
                  sample = FALSE)

```

