---
title: "The state of UK public sector analysis code: 2026"
output:
  html:
    self-contained: true
---

```{r echo=FALSE}

data <- CARS::get_tidy_data_file("2026_data.csv")
config <- CARS::load_configs("../../configs/")

waves_data <- CARS::get_all_waves()

data <- waves_data$"2026"

all_wave_data <- dplyr::bind_rows(waves_data)

tables <- CARS::summarise_all(data, config, all_tables = TRUE)

```

## How to use this research

Responding to CARS is voluntary. The results presented here are from a self-selecting sample of government analysts. Because respondents are self-selecting, the results we present reflect the views of the analysts who participated.

For more detail, [see the data collection page](data_collection.qmd).

# Coding frequency and tools

We asked respondents if they had any coding experience, inside or outside of work. Of `r nrow(data)` respondents, `r round(sum(data$coding_exp == "Yes") / nrow(data) * 100, 1)`% had coding experience. We asked all respondents with coding experience "In your current role, how often do you write code to complete your work objectives?"

To see how this compares to other years, [see the data collection page](data_collection.qmd#coding-frequency)

## Coding frequency

#### 2026 data

```{r}

plot <- CARS::plot_freqs(data = tables$code_freq, 
                   xlab = "How often do you write code in your current role?",
                   font_size = 14)

table <- CARS::df_to_table(data = tables, 
                           config = config, 
                           question = "code_freq")

CARS::wrap_outputs("coding-freq", plot, table)

```

## Coding interest 

In 2026, we also asked respondents with coding experience whether they would like to use more coding in their role.

```{r}

plot <- CARS::plot_freqs(data = tables$coding_freq_pref, 
                   xlab = "Would you prefer to use more coding in your role?",
                   font_size = 14,
                   type = "pie")

table <- CARS::df_to_table(data = tables, 
                           config = config, 
                           question = "coding_freq_pref")

CARS::wrap_outputs("coding-freq-pref", plot, table)

```

#### Knowledge and use of programming languages

### Knowledge of coding tools

Given a list of programming tools, we asked all respondents with coding experience whether they knew how to use each tool, regardless of whether they currently use it in their work. Note this includes respondents who do not code in their current role.

Please note that capability in programming languages is self-reported here and was not objectively defined or tested. 

```{r}
tables$knowledge <- tables$knowledge |>
  dplyr::arrange(dplyr::desc(value == "Yes"), dplyr::desc(n))

plot <- CARS::plot_stacked(tables$knowledge,  xlab = "Programming tool", colour_scale = "scale", font_size = 14)
table <- CARS::df_to_table(tables$knowledge,  column_headers = c("Programming tool", "Yes", "No"), crosstab = TRUE)

CARS::wrap_outputs("knowledge", plot, table)

```

### Use of coding tools

Using the same list of programming tools, we asked respondents which tools they currently use in their role.



```{r}
tables$use <- tables$use |>
  dplyr::arrange(dplyr::desc(value == "Yes"), dplyr::desc(n))

plot <- CARS::plot_stacked(tables$use, xlab = "Programming tool", colour_scale = "scale", font_size = 14)
table <- CARS::df_to_table(tables$use, column_headers = c("Programming tool", "Yes", "No"), crosstab = TRUE)

CARS::wrap_outputs("use", plot, table)

```



### Use compared with knowledge of coding tools

The chart below compares the proportion of respondents with coding experience who reported knowing how to use each programming tool with the proportion who reported using it in their current role.

```{r}

tools_compare <- dplyr::bind_rows(
  tables$use |>
    dplyr::filter(value == "Yes") |>
    dplyr::transmute(name, status = "Use", n, count, sample),
  tables$knowledge |>
    dplyr::filter(value == "Yes") |>
    dplyr::transmute(name, status = "Knowledge", n, count, sample)
)

tool_order <- tools_compare |>
  dplyr::group_by(name) |>
  dplyr::summarise(total = sum(n, na.rm = TRUE), .groups = "drop") |>
  dplyr::arrange(dplyr::desc(total)) |>
  dplyr::pull(name)

tools_compare <- tools_compare |>
  dplyr::mutate(
    name = factor(name, levels = tool_order),
    status = factor(status, levels = c("Use", "Knowledge"))
  ) |>
  dplyr::arrange(name, status)

plot <- CARS::plot_dumbbell(tools_compare,
                            group_order = c("Use", "Knowledge"),
                            xlab = "Percent",
                            ylab = "Programming tool",
                            font_size = 14)

table <- CARS::df_to_table(tools_compare[c(1:3)],
                           crosstab = TRUE,
                           column_headers = c("Programming tool", "Use", "Knowledge"),
                           sample = FALSE)

CARS::wrap_outputs("tools-compare", plot, table)

```

### Open source capability over time

The proportion of respondents who report having the capability to use R and Python, is shown alongside the proportion who are able to use SAS, SPSS or Stata, for the past four years of the survey.

```{r}

freqs <- CARS::summarise_os_vs_prop(all_wave_data)

colnames(freqs) <- gsub("percent", "n", colnames(freqs))
freqs <- dplyr::relocate(freqs, year)
freqs <- dplyr::relocate(freqs, n, .after = "lang_type")

table <- CARS::df_to_table(freqs[c(1:3, 6, 7)],
                           column_headers = c("Programming language type",
                                              "Year",
                                              "Know how to programme with these tools (percent)",
                                              "Lower confidence limit (percent)",
                                              "Upper confidence limit (percent)"), 
                           sample = FALSE)



plot <- CARS::plot_stacked(freqs[c(1:4,6)], 
                           type = "line", 
                           orientation = "v", 
                           font_size = 14,
                           ylab = "Know how to programme with these tools", 
                           xlab = "Year",
                           error_y = CARS::set_error_bars(freqs$lower_ci, (freqs$upper_ci))) |>
  CARS::set_axis_range(0, 1, axis = "y") |>
  plotly::layout(legend = list(x = 100, y = 0.8))

CARS::wrap_outputs("tools-over-time", plot, table)

```

### Professions capability in different tools (%)

Differences in preferred languages may lead to silos between analytical professions. Here we show the percentage of respondents reporting capability in different tools, within the different analytical professions.

Please note that respondents might be members of more than one profession, and may report capability in more than one tool.

```{R}

CARS::df_to_table(tables$languages_by_prof[c(2,1,3)], 
            crosstab = TRUE, 
            sample = FALSE,
            heatmap = TRUE,
            show_percent_symbol = FALSE)
```

### Git

We asked respondents to answer "Yes", "No" for the following question:

-   Do you know how to use Git?

Please note these outputs include people with coding experience who do not code in their current role.

```{r}
plot <- CARS::plot_freqs(data = tables$git, 
                         font_size = 14, 
                         xlab = "Do you know how to use Git?",
                         type = "pie")

table <- CARS::df_to_table(data = tables$git, 
                           column_headers = c("Do you know how to use Git?", 
                                              "Percentage"))

CARS::wrap_outputs("git", plot, table)

```

## Coding capability and change

### Coding experience

We asked respondents with coding experience how many years of experience they had in a coding role, excluding any years in education.

This data includes any experience in other roles and sectors, but does not define the type of experience.

```{r}

plot <- CARS::plot_freqs(tables$coding_years,
                         xlab = "How many years of experience do you have in a coding role?", 
                         font_size = 14,
                         orientation = "h")
table <- CARS::df_to_table(data = tables,
                           config = config,
                           question = "coding_years")

CARS::wrap_outputs("coding-years", plot, table)

```

### Change in coding ability during current role

We asked "Has your coding ability changed during your current role?"

This question was only asked of respondents with coding experience outside of their current role. This means analysts who first learned to code in their current role are not included in the data.

```{r}

plot <- CARS::plot_freqs(data = tables,
                   config = config,
                   question = "ability_change",
                   xlab = "How has your coding ability changed during your current role?",
                   font_size = 14,
                   orientation = "h")

table <- CARS::df_to_table(data = tables,
                           config = config,
                           question = "ability_change")

CARS::wrap_outputs("ability-change", plot, table)

```

### Managing a coding project

We asked respondents if they'd feel confident managing a coding project.

```{r}

plot <- CARS::plot_freqs(data = tables$manage_project, 
                   xlab = "Would you feel confident managing a coding project?",
                   font_size = 14,
                   type = "pie")

table <- CARS::df_to_table(data = tables, 
                           config = config, 
                           question = "manage_project")

CARS::wrap_outputs("manage-project", plot, table)

```

## Motivation and barriers to coding

### Motivation

Respondents who do not currently code in their role were asked whether they would like to use coding as part of their work.

```{r}

plot <- CARS::plot_freqs(data = tables,
                   config = config,
                   question = "coding_learn_pref",
                   xlab = "Would you like to use coding in your current role?",
                   font_size = 14,
                   type = "pie")

table <- CARS::df_to_table(data = tables,
                           config = config,
                           question = "coding_learn_pref")

CARS::wrap_outputs("coding-learn-pref", plot, table)

```

### Specific barriers

We asked respondents about any barriers they face in learning to code in their current role. The most commonly cited barrier was a lack of opportunity.

```{r}

coding_improve_barriers <- tables$coding_improve_barriers |>
  dplyr::filter(value == "Yes") |>
  dplyr::filter(!(grepl("^Other", name, ignore.case = TRUE) & n == 0)) |>
  dplyr::arrange(desc(n))

plot <- CARS::plot_stacked(coding_improve_barriers,  
                           xlab = "Barrier", 
                           colour_scale = "gradient", 
                           font_size = 14) |>
  plotly::layout(showlegend = FALSE)

table <- CARS::df_to_table(coding_improve_barriers,  
                           column_headers = c("Barrier", "Yes"), 
                           crosstab = TRUE)

CARS::wrap_outputs("coding-improve-barriers", plot, table)

```

We also asked respondents about any barriers to improving their coding skills in their current role.

```{r}

coding_learn_barriers <- tables$coding_learn_barriers |>
  dplyr::filter(value == "Yes") |>
  dplyr::filter(!(grepl("^Other", name, ignore.case = TRUE) & n == 0)) |>
  dplyr::arrange(desc(n))

plot <- CARS::plot_stacked(coding_learn_barriers,  
                           xlab = "Barrier", 
                           colour_scale = "gradient", 
                           font_size = 14) |>
  plotly::layout(showlegend = FALSE)

table <- CARS::df_to_table(coding_learn_barriers,  
                           column_headers = c("Barrier", "Yes"), 
                           crosstab = TRUE)

CARS::wrap_outputs("coding-learn-barriers", plot, table)

```

## AI tools

### Use of AI tools

We asked all respondents who use code in their role, it they use AI tools when coding at work.

The term 'AI' was not defined in this question.

```{r}
plot <- CARS::plot_freqs(data = tables,
                   config = config,
                   question = "ai",
                   xlab = "Do you use any AI assistants when coding at work?",
                   font_size = 14,
                   type = "pie")

table <- CARS::df_to_table(data = tables,
                           config = config,
                           question = "ai")

CARS::wrap_outputs("ai", plot, table)
```

### AI tool use by profession

Overall, `r round(tables$ai[tables$ai[[1]] == "Yes", "n"][[1]] * 100, 1)`% of coders reported using AI tools when coding at work. The chart below breaks this down by analytical profession, showing the proportion of coders who use AI assistants.

```{r}

prof_q <- CARS::get_question_data(config, "professions")
prof_cols <- prof_q$cols[!grepl("none|other", prof_q$cols)]
prof_labels <- config[["professions"]][["cols"]]
prof_labels <- prof_labels[!grepl("no profession|Other", prof_labels)]

ai_by_prof <- lapply(prof_cols, function(col) {
  filtered_data <- dplyr::filter(data, get(col) == "Yes")

  if (nrow(filtered_data) > 0) {
    ai_freq <- CARS::summarise_data(filtered_data, config, question = "ai", sample = TRUE)
    ai_yes <- ai_freq[ai_freq[[1]] == "Yes", , drop = FALSE]

    if (nrow(ai_yes) > 0) {
      ai_yes$profession_col <- col
      return(ai_yes)
    }
  }
})

ai_by_prof <- do.call(rbind, ai_by_prof)

ai_by_prof <- ai_by_prof |>
  dplyr::transmute(
    name = dplyr::recode(profession_col, !!!prof_labels),
    n = n,
    count = count,
    sample = sample
  ) |>
  dplyr::arrange(dplyr::desc(n))

ai_by_prof_plot <- ai_by_prof
colnames(ai_by_prof_plot)[colnames(ai_by_prof_plot) == "sample"] <- "denominator"

plot <- CARS::plot_freqs(ai_by_prof_plot,
                         xlab = "Profession",
                         font_size = 14,
                         orientation = "h")

table <- CARS::df_to_table(ai_by_prof,
                           column_headers = c("Profession", "Yes"),
                           sample = FALSE)

CARS::wrap_outputs("ai-by-prof", plot, table)
```

### How AI tools are used

We asked all respondents who use AI tools when coding at work, what they use these tools for.

```{r}

ai_use_data <- dplyr::filter(tables$ai_use, name != "Other") 

plot <- CARS::plot_likert(ai_use_data,
                          mid = 3, neutral_mid = TRUE,
                          break_q_names_col = "name",
                          max_lines = 2, font_size = 14,
                          height = 800, width = "100%")

table <- CARS::df_to_table(ai_use_data,
                           column_headers = c("AI use", "Never", "Rarely", "Sometimes", "Regularly", "Always"), crosstab = TRUE)


CARS::wrap_outputs("ai-use", plot, table)
```

## Cloud tools

### Use of cloud tools

Respondents were asked if they use any cloud data platforms in their current role.

```{r}

cloud <- tables$cloud |>
  dplyr::filter(value == "Yes") |>
  dplyr::filter(!(grepl("^Other", name, ignore.case = TRUE) & n == 0)) |>
  dplyr::arrange(desc(n))

plot <- CARS::plot_stacked(cloud,  
                           xlab = "Data Platform", 
                           colour_scale = "gradient", 
                           font_size = 14) |>
  plotly::layout(showlegend = FALSE)

table <- CARS::df_to_table(cloud,  
                           column_headers = c("Data Platform", "Yes"), 
                           crosstab = TRUE)

CARS::wrap_outputs("cloud", plot, table)

```

## Reproducible analytical pipelines (RAP)

RAP refers to the use of good software engineering practices to make analysis pipelines more reproducible. This approach aims to use automation to improve the quality and efficiency of analytical processes.

The following links contain more resources on RAP:

-   you can find minimum RAP standards in the [RAP MVP](rap_mvp_maturity_guidance/Reproducible-Analytical-Pipelines-MVP.md%20at%20master%20·%20best-practice-and-impact/rap_mvp_maturity_guidance%20·%20GitHub)
-   you can find guidance on quality assuring code in the [Duck Book](https://best-practice-and-impact.github.io/qa-of-code-guidance/intro.html)

### Implementing RAP

We asked respondents to what extent they agreed with the statement: "I am currently implementing RAP in my work."

```{r}

plot <- CARS::plot_freqs(data = tables$rap_implementing, 
                   xlab = "'I am currently implementing RAP in my work'",
                   font_size = 14)

table <- CARS::df_to_table(data = tables, 
                           config = config, 
                           question = "rap_implementing")

CARS::wrap_outputs("rap-implementing", plot, table)

```

### RAP Barriers

We also asked respondents about the main barriers to implementing RAP in their current role. The most commonly reported barrier was a lack of time.

```{r}

rap_barriers <- tables$rap_barriers |>
  dplyr::filter(value == "Yes") |>
  dplyr::filter(!(grepl("^Other", name, ignore.case = TRUE) & n == 0)) |>
  dplyr::arrange(desc(n))

plot <- CARS::plot_stacked(rap_barriers,  
                           xlab = "Barrier", 
                           colour_scale = "gradient", 
                           font_size = 14) |>
  plotly::layout(showlegend = FALSE)

table <- CARS::df_to_table(rap_barriers,  
                           column_headers = c("Barrier", "Yes"), 
                           crosstab = TRUE)

CARS::wrap_outputs("rap-barriers", plot, table)

```

## Good coding practices

We asked respondents who reported writing code at work about the good practices they apply when writing code at work. These questions cover many of the coding practices recommended in the quality assurance of code for analysis and research guidance, as well as the [minimum RAP standards](https://github.com/best-practice-and-impact/rap_mvp_maturity_guidance/blob/master/Reproducible-Analytical-Pipelines-MVP.md) set by the cross-government [RAP champions network](https://analysisfunction.civilservice.gov.uk/support/reproducible-analytical-pipelines/reproducible-analytical-pipeline-rap-champions/).

Coding practices have been classified as either 'Basic' or 'Advanced'. Basic practices are those that make up the minimum RAP standards, while Advanced practices help improve reproducibility. The percentage of respondents who reported applying these practices either 'Regularly' or 'All the time' is shown below.

Open sourcing was defined as 'making code freely available to be modified and redistributed'

```{r}

tables$rap_components <-  tables$rap_components |> dplyr::arrange(dplyr::desc(value), dplyr::desc(n)) |> data.frame()
 
tables$rap_components$name <- as.character(tables$rap_components$name)
# 
tables$rap_components$name <- factor(tables$rap_components$name, levels = tables$rap_components$name)

plot <- CARS::plot_grouped(tables$rap_components,
                           orientation = "h",
                           xlab = "RAP component", font_size = 14)

table <- CARS::df_to_table(tables$rap_components,
                           column_headers = c("RAP component",
                                              "Type",
                                              "Percentage of analysts who code in their work"))

CARS::wrap_outputs("rap-comp", plot, table)

```

### RAP components over time

The table below shows RAP component use over time for 2023, 2024 and 2026. Empty cells indicate that a component was not asked in that year.

```{r}

years <- c("2023", "2024", "2026")

rap_labels <- unlist(config[["rap_components"]][["cols"]], use.names = TRUE)

rap_components_by_year <- lapply(years, function(year) {
  year_data <- waves_data[[year]]
  available_cols <- intersect(names(rap_labels), colnames(year_data))

  if (length(available_cols) == 0) {
    return(NULL)
  }

  denominator <- sum(year_data$code_freq != "Never", na.rm = TRUE)

  if (denominator == 0) {
    return(data.frame(
      name = unname(rap_labels[available_cols]),
      year = year,
      n = NA_real_,
      stringsAsFactors = FALSE
    ))
  }

  data.frame(
    name = unname(rap_labels[available_cols]),
    year = year,
    n = colSums(year_data[available_cols], na.rm = TRUE) / denominator,
    stringsAsFactors = FALSE
  )
}) |>
  dplyr::bind_rows()

rap_components_over_time <- data.frame(
  name = unname(rap_labels),
  year = rep(years, each = length(rap_labels)),
  stringsAsFactors = FALSE
) |>
  dplyr::left_join(rap_components_by_year, by = c("name", "year"))

component_order <- rap_components_over_time |>
  dplyr::filter(year == "2026") |>
  dplyr::arrange(dplyr::desc(n)) |>
  dplyr::pull(name)

if (length(component_order) == 0) {
  component_order <- unname(rap_labels)
}

rap_components_over_time <- rap_components_over_time |>
  dplyr::mutate(
    name = factor(name, levels = component_order),
    year = factor(year, levels = years)
  ) |>
  dplyr::arrange(name, year)

CARS::df_to_table(
  rap_components_over_time[c("name", "year", "n")],
  crosstab = TRUE,
  sample = FALSE,
  heatmap = TRUE,
  crosstab_global_scale = TRUE,
  show_percent_symbol = FALSE,
  column_headers = c("RAP component", years)
)

```

### Consistency of good coding practices

We asked respondents who reported writing code at work how frequently they apply good coding practices when writing code at work.

```{r}

plot <- CARS::plot_likert(tables$coding_practices,
                          mid = 5, neutral_mid = FALSE,
                          break_q_names_col = "name",
                          max_lines = 2, font_size = 14,
                          height = 800, width = "100%")

table <- CARS::df_to_table(tables$coding_practices,
                           column_headers = c(
                             "Statement",
                             "I don't understand this question (%)",
                             "Never (%)",
                             "Rarely (%)",
                             "Sometimes (%)",
                             "Regularly (%)",
                             "Always (%)"
                             ),
                           crosstab = TRUE)

CARS::wrap_outputs("good-practices", plot, table)

```

### Code documentation

We asked respondents who reported writing code at work how frequently they write different forms of documentation when programming in their current role.

[Embedded documentation](https://best-practice-and-impact.github.io/qa-of-code-guidance/code_documentation.html) is one of the components which make up a RAP minimum viable product. Documentation is important to help others be clear on how to use the product and what the code is intended to do.

```{r}

plot <- CARS::plot_likert(tables$doc,
                          mid = 5, neutral_mid = FALSE,
                          break_q_names_col = "name",
                          max_lines = 2, font_size = 14,
                          height = 800, width = "100%")

table <- CARS::df_to_table(tables$doc,
                           column_headers = c(
                             "Statement",
                             "I don't understand this question (%)",
                             "Never (%)",
                             "Rarely (%)",
                             "Sometimes (%)",
                             "Regularly (%)",
                             "Always (%)"
                             ),
                           crosstab = TRUE)

CARS::wrap_outputs("doc", plot, table)

```

### Working practices

We asked respondents who reported writing code at work how frequently they use good working practices in the coding projects they work on.

```{r}

plot <- CARS::plot_likert(tables$working_practices,
                          mid = 5, neutral_mid = FALSE,
                          break_q_names_col = "name",
                          max_lines = 2, font_size = 14,
                          height = 800, width = "100%")

table <- CARS::df_to_table(tables$working_practices,
                           column_headers = c(
                             "Statement",
                             "I don't understand this question (%)",
                             "Never (%)",
                             "Rarely (%)",
                             "Sometimes (%)",
                             "Regularly (%)",
                             "Always (%)"
                             ),
                           crosstab = TRUE)

CARS::wrap_outputs("working_practices", plot, table)

```

### Standards

Respondents were asked whether they follow any analysis or coding quality standards, whether that be either internal standards, external guidance, or a combination of both.

```{r}

plot <- CARS::plot_freqs(data = tables$standards, 
                   xlab = "Do you work to any analysis or coding quality standards?",
                   font_size = 14,
                   orientation = "h")

table <- CARS::df_to_table(data = tables, 
                           config = config, 
                           question = "standards")

CARS::wrap_outputs("standards", plot, table)

```

### Duck Book

Respondents were asked whether they were aware of the Duck Book.

The [Duck Book](https://best-practice-and-impact.github.io/qa-of-code-guidance/intro.html) is guidance produced by the ONS Analysis Standards and Pipelines Hub. It outlines software engineering best practices tailored to analysts and researchers who work with data using code.

```{r}

plot <- CARS::plot_freqs(data = tables$duck_book, 
                   xlab = "Are you aware of the 'Duck Book'?",
                   font_size = 14,
                   type = "pie")

table <- CARS::df_to_table(data = tables, 
                           config = config, 
                           question = "duck_book")

CARS::wrap_outputs("duck_book", plot, table)

```

### Packages

In 2026, we asked respondents whether they or their team use cross-government or department-specific analysis packages in their work.

See the list of packages reported by respondents on the [useful pages page](useful_packages.qmd).

```{r}

plot <- CARS::plot_freqs(data = tables$packages, 
                   xlab = "Do you or your team use any cross-government or department-specific analysis packages in your work?",
                   font_size = 14,
                   type = "pie")

table <- CARS::df_to_table(data = tables, 
                           config = config, 
                           question = "packages")

CARS::wrap_outputs("packages", plot, table)

```
